\documentclass[10pt]{article}
\usepackage[margin=0.72in,top=0.6in,bottom=0.6in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{titlesec}

\setstretch{1.0}
\tolerance=400
\emergencystretch=1em
\setlength{\parskip}{2pt plus 1pt}
\setlength{\parindent}{0.7em}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=blue}
\setlist{nosep,leftmargin=1.3em,topsep=2pt,partopsep=0pt}
\titlespacing*{\section}{0pt}{6pt plus 2pt}{3pt}
\titlespacing*{\subsection}{0pt}{4pt plus 1pt}{2pt}
\titleformat{\section}{\normalfont\bfseries\large}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalfont\bfseries\normalsize}{\thesubsection}{0.5em}{}

\lstset{
  basicstyle=\ttfamily\scriptsize,
  keywordstyle=\bfseries,
  commentstyle=\color{gray},
  breaklines=true,
  frame=single,
  framesep=2pt,
  xleftmargin=3pt,
  xrightmargin=3pt,
  aboveskip=3pt,
  belowskip=3pt,
}

\title{\vspace{-2.0em}Task 3 -- Prototype Implementation: MongoDB + AirBnB\vspace{-0.4em}}
\author{Abhinav Bachu, Hanshul Bahl, Alejandra Arias\\CS 498 / Data Management in the Cloud}
\date{04/20/2026}

\begin{document}
\maketitle
\vspace{-2.0em}

\section{Stage 3 Goal and Scope}

Stage~2 designed four collections (\texttt{listings}, \texttt{calendar}, \texttt{reviews}, \texttt{neighborhoods}), eight compound indexes, and pseudocode for all six AirBnB queries. Stage~3 turns that design into a working prototype that boots MongoDB via \texttt{mongodb-memory-server}, loads real InsideAirbnb data for four cities (Los Angeles, Portland, Salem, San Diego), and runs three of the six queries end-to-end with \texttt{explain('executionStats')} captured so we can show \emph{actual} index usage rather than promised index usage. The schema and indexes are unchanged from Stage~2~\S2.2; everything else---loader, queries, driver, evidence---is new in \texttt{stage3-prototype/}. The prototype runs with one \texttt{npm~start} in under 10\,s on a laptop, no external Mongo required.

\section{Implemented Subset (3 of 6 Queries)}

We picked Q1, Q5, and Q6 because together they cover the three distinct query \emph{styles}. \textbf{Q1} (Portland 2-day search) demonstrates filter\,+\,group\,+\,lookup\,+\,sort and exercises both the \texttt{calendar} date index and the \texttt{listings} unique key. \textbf{Q5} (December reviews per city per year) is a pure aggregation over \texttt{reviews}. \textbf{Q6} (re-book reminder + same-host listings) is multi-step orchestration touching all four collections. We skipped Q3/Q4 because both rely on the same run-segmentation logic Stage~2 flagged as awkward in pure aggregation, and Q2 is essentially a set difference on the same \texttt{calendar} index pattern as Q1.

\section{Loader Architecture and Data Volume}

\texttt{src/a.js} streams \texttt{listings.csv.gz}, \texttt{reviews.csv.gz}, \texttt{calendar.csv.gz}, and \texttt{neighbourhoods.csv} from \texttt{data.insideairbnb.com} through \texttt{zlib.createGunzip()} into \texttt{csv-parse} (no full CSV ever held in memory), cleans rows (currency strip, \texttt{t/f} to boolean, neighborhood dedup), bulk-inserts, and creates the eight Stage~2 indexes. Recorded run: \textbf{1{,}780 listings, 4{,}000 reviews, 411 neighborhoods, 8{,}000 calendar rows}---enough for every query to return non-empty results, small enough to iterate on queries in seconds. Evidence at \texttt{stage3-prototype/out/load\_evidence.txt}.

\section{Query Implementations and Execution Plans}

\subsection{Q1 -- Portland 2-day search}
\begin{lstlisting}
[ {$match: {city:"portland", date:{$in:dates}, available:true}},
  {$group: {_id:"$listing_id", a:{$sum:1}}}, {$match:{a:dates.length}},
  {$lookup:{from:"listings", let:{l:"$_id"}, as:"b", pipeline:[
     {$match:{$expr:{$and:[{$eq:["$city","portland"]},{$eq:["$listing_id","$$l"]}]}}},
     {$project:{_id:0,name:1,neighborhood:1,room_type:1,accommodates:1,property_type:1,price:1,review_scores_rating:1}}]}},
  {$unwind:"$b"},{$replaceRoot:{newRoot:"$b"}},
  {$sort:{review_scores_rating:-1, listing_id:1}}, {$limit:25} ]
\end{lstlisting}
The driver picks the demo dates dynamically by querying for the earliest pair of consecutive available dates in Portland, so the query always has live data to operate on. \textbf{Plan} (from \texttt{out/explain\_q1.json}): the leading \texttt{\$match} hits \texttt{IXSCAN} on \texttt{city\_1\_date\_1\_available\_1} with bounds locked to Portland and the two requested dates. \texttt{nReturned=3, totalKeysExamined=6, totalDocsExamined=6, executionTimeMillis=3}---one key and one document per matching \texttt{(listing,~date)} pair. The \texttt{\$lookup} hits \texttt{city\_1\_listing\_id\_1} (unique) for 3 doc reads. The \texttt{\$sort} runs over 3 docs with no spill. Top result in the recorded run: ``Beautiful Apt near City Center!'' (Irvington, rating 4.94) for window 2025-12-04~/~2025-12-05.

\subsection{Q5 -- December reviews per city per year}
\begin{lstlisting}
[ {$match: {date: {$regex: /^\d{4}-12-\d{2}$/}}},
  {$group: {_id:{city:"$city", year:{$substr:["$date",0,4]}}, review_count:{$sum:1}}},
  {$project:{_id:0, city:"$_id.city", year:"$_id.year", review_count:1}},
  {$sort:{city:1, year:1}} ]
\end{lstlisting}
Returned 53~(city,~year) groups in 2~ms over 4{,}000 docs in the recorded run. This is a full-collection aggregation by design---every December review must be counted---so no useful index prefix applies. The Stage~2 design intentionally did not include a leading \texttt{\{date:1\}} index on \texttt{reviews} because Q5 is the only date-driven \texttt{reviews} query and it already runs in 2~ms. If \texttt{reviews} grew to 10M+ documents, adding \texttt{reviews~\{date:1\}} would be the first tuning step.

\subsection{Q6 -- Re-book reminder + same-host listings}
\textbf{Step~1} (one aggregation): group \texttt{reviews} by \texttt{(city, listing\_id, reviewer\_id)}, keep only groups with \texttt{review\_count$>$1}. \textbf{Step~2} (per group, in app code, capped at 25 records for the demo): for each prior review date, compute the month bounds, then run one \texttt{findOne} on \texttt{calendar} for in-month availability, one aggregate on \texttt{calendar} for min/max nights in that month, one \texttt{findOne} on \texttt{listings} for description and host metadata, and one \texttt{find} on \texttt{listings} for the host's other listings in the same city. Every step-2 read lands on an index already designed in Stage~2: \texttt{city\_1\_listing\_id\_1\_date\_1} for calendar, \texttt{city\_1\_listing\_id\_1} (unique) and \texttt{host\_id\_1\_city\_1} for listings. \textbf{We did not have to define a single new index for Q6}, which is itself a validation of the Stage~2 index design. Recorded run produced 25 enriched records, including repeat reviews for ``RUN Runyon, Beau Furn Rms Terrace Hollyw Hill View'' (host \emph{Chas.}) with one other same-host listing in Los Angeles. Decomposing this query into one aggregation plus a small handful of indexed point reads---rather than one giant nested-\texttt{\$lookup} pipeline---kept the code straight-line and easy to debug, and incurred essentially no runtime cost because every read hits an index.

\section{Critique: What We'd Change}

(1)~\textbf{Date-as-string was the right call but for the wrong stated reason.} Stage~2 justified storing dates as ISO strings primarily on storage grounds. In practice, the bigger payoff turned out to be \emph{query simplicity}: Q5's \texttt{\$substr} extraction stays trivial and is regex-prefix-indexable, whereas \texttt{\$year}/\texttt{\$month} on a true \texttt{Date} forces a per-document compute that no index can short-circuit. We would re-frame the trade-off in Stage~2's design rationale.

(2)~\textbf{The \texttt{host\_id\_1\_city\_1} index has the wrong key order.} Stage~2 chose this for Q6's same-host lookup. In actual use the lookup also filters by \texttt{listing\_id$\neq$current}, and cardinality is dominated by \texttt{host\_id} (most hosts have 1--2 listings). Reversing it to \texttt{\{city, host\_id\}} would let the index share a \texttt{city} prefix with our three other listings indexes (\texttt{\{city, listing\_id\}}, \texttt{\{city, neighborhood, room\_type\}}), which is more consistent with the rest of the design. The current order is correct, just slightly off-pattern.

(3)~\textbf{\texttt{neighborhoods} could be embedded as a per-city array.} It is essentially a small read-only join lookup table. If we ever needed Q2 frequently, promoting the canonical neighborhood list into a one-document-per-city ``city metadata'' shape would let us skip the \texttt{\$lookup} entirely.

\section{Lessons Learned and Advice}

\textbf{Lessons.} \texttt{mongodb-memory-server} shrinks the design loop to seconds and should be the default for prototyping. \texttt{explain('executionStats')} is the single most valuable design-validation tool we used---one line of \texttt{.explain()} per query turned guessed index usage into proven index usage. Multi-step orchestration in application code (Q6) is meaningfully easier to read and debug than a single deeply-nested \texttt{\$lookup} aggregation, and at our scale costs essentially nothing because each per-record read still hits an index. Schema design happens at \emph{index}-design time, not at collection-design time: every query's actual performance was determined by compound-index choices, not by document shape.

\textbf{Advice.} Pick one query first and design every collection and every index around it; only then add the second. Doing all six in one pass tempts you into ``just-in-case'' indexes (we did this with \texttt{host\_id\_1\_city\_1}). Start with \texttt{mongodb-memory-server}, not Atlas---switching is one URI change. Always commit \texttt{explain} output alongside your query results so reviewers (and future you) can verify what the planner is actually doing. Resist the temptation to denormalize early: we considered embedding \texttt{calendar} arrays inside \texttt{listings} and the cost of being wrong about that compounds quickly.

\section{My Role in the Project}

I led infrastructure and data ingestion for the prototype. Concretely: I scaffolded the \texttt{stage3-prototype/} project, picked \texttt{mongodb-memory-server} so the team could iterate on queries without an Atlas account, and wrote the loader (\texttt{src/a.js})---streaming gzipped CSVs from \texttt{data.insideairbnb.com} through \texttt{zlib} and \texttt{csv-parse}, normalizing types, deduplicating neighborhoods to satisfy the unique index, and bulk-inserting into the four collections with all eight Stage~2 indexes built right after. I also wrote the driver (\texttt{src/e.js}) that boots the in-process Mongo, calls the loader, runs all three queries, captures the \texttt{explain('executionStats')} output, and writes everything in \texttt{out/} so the evidence is reproducible. On the query side, I implemented Q1 (Portland 2-day search) including the dynamic date-picking helper that finds the earliest pair of consecutive available Portland dates so the demo always returns live results. Outside of code, I owned the GitHub repository (\texttt{cs498-airbnb-mongodb-prototype}), wrote the top-level \texttt{README}, and built the PDF generation pipeline (\texttt{tectonic} for the report, \texttt{md-to-pdf} for the slide outline). The biggest single decision I argued for was \emph{not} setting up Atlas in week~1: spinning up a real cluster early would have cost us several days of glue work, while the in-process server let us focus on schema and query correctness from day one. Throughout, I worked closely with my teammates: I integrated their query implementations into the driver and re-ran the evidence pipeline whenever the schema or query logic changed.

\newpage
\section{References}

\begin{enumerate}[label={[\arabic*]}]
\item MongoDB Documentation. \url{https://www.mongodb.com/docs/manual/}
\item MongoDB Aggregation Pipeline. \url{https://www.mongodb.com/docs/manual/core/aggregation-pipeline/}
\item MongoDB Indexes. \url{https://www.mongodb.com/docs/manual/indexes/}
\item MongoDB \texttt{cursor.explain()} and \texttt{executionStats}. \url{https://www.mongodb.com/docs/manual/reference/method/cursor.explain/}
\item MongoDB \texttt{\$lookup}. \url{https://www.mongodb.com/docs/manual/reference/operator/aggregation/lookup/}
\item MongoDB Data Modeling. \url{https://www.mongodb.com/docs/manual/core/data-model-design/}
\item \texttt{mongodb-memory-server}. \url{https://github.com/typegoose/mongodb-memory-server}
\item \texttt{csv-parse}. \url{https://csv.js.org/parse/}
\item InsideAirbnb Data Portal. \url{https://insideairbnb.com/get-the-data.html}
\item AirBnB Assignment Query Document. \texttt{AirBnB-info-questions.pdf}.
\item Stage~2 Report. \texttt{task3-deliverables/TASK\_3\_Report\_MongoDB\_Airbnb.pdf}.
\item Project repository. \url{https://github.com/abachu2005/cs498-airbnb-mongodb-prototype}
\end{enumerate}

\end{document}
